---
title: "Conditional/ControlFlow Notebook"
output: html_notebook
---

#### Basic Comparisons

The comparisons are self explanatory:

```{r}
# Comparison of logicals
TRUE == FALSE

# Comparison of numerics
-6 * 14 != 17 - 101

# Comparison of character strings
"useR" == "user"

# Compare a logical with a numeric
TRUE == 1
```

Note: Remember about TRUE == 1 and FALSE == 0 being TRUE because of falsey and truthy comparisons.  


#### Now some weird examples:

When comparing strings, comparison is determined by alphabetic order.

As described in the previous subsection, TRUE corresponds to 1 and FALSE corresponds to 0 in R. 

```{r}
(1 + 2) > 4

"dog" < "Cats"

TRUE <= FALSE
```

#### Comparing two vectors

Suppose we have two vectors for no. of views in two different social media platforms.  Let us perform some basic comparison as described in the in-line script below:

```{r}
# The linkedin and facebook vectors were declared
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# On which days did the number of LinkedIn profile views exceed 15?
linkedin > 15

# When was your LinkedIn profile viewed only 5 times or fewer?
linkedin <= 5

# LinkedIn more popular than Facebook
linkedin > facebook
```

#### Comparing all values inside a matrix 

```{r}
# The social data has been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)
views <- matrix(c(linkedin, facebook), nrow = 2, byrow = TRUE)

# When does views equal 13?  
views  == 13

# When is views less than or equal to 14?
views <= 14
```

#### & and |

Unlike in other languages I know, R utilizes a single & and | for basic comparison.  Here are some basic uses and examples for & and |.

```{r}
# The linkedin and last variable are already defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
last <- tail(linkedin, 1)

# Is last under 5 or above 10?
last < 5 | last > 10

# Is last between 15 (exclusive) and 20 (inclusive)?
last > 15 & last <= 20
```

#### Using sum() in counting TRUEs in an vector

Remember that TRUE is counted as 1 in R. As an example:

```{r}
# declaring a logical vector
logical_vector <- c(TRUE, FALSE, TRUE)
# count TRUES
sum(logical_vector)
```

As expected, `sum()` was able to tally the number of TRUE as 2.  

#### if statement

if statement is practically the same in other programming languages.  Example is below:  

```{r}
# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Examine the if statement for medium
if(medium == "LinkedIn"){
  print("Showing LinkedIn information")
}

# Write the if statement for num_views
if (num_views > 15) {
  print("You're popular!")
} else {
  print("Not POPULAR")
}
```

Another similar example with a bunch of else if statement is below:

```{r}
# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Control structure for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
} else if (medium == "Facebook") {
  print("Showing Facebook information")
} else {
  print("Unknown medium")
}

# Control structure for num_views
if (num_views > 15) {
  print("You're popular!")
} else if (num_views <= 15 & num_views > 10) {
  print("Your number of views is average")
} else {
  print("Try to be more visible!")
}
```

