---
title: "Apply Family"
output: html_notebook
---

#### Basics of Apply

Apply allows us to simplify tasks by applying all items in a list/vector/matrix to a function without having to write a complicated for loop statement to iterate over the data structure we want to analyze.

#### Basics on apply

There's a family of apply function but let us examine `lapply()` and using it to split four data by cutting the semi-colon and making all entries small letters.

Again, we need to use `strsplit()` to split our character data and eventually need to use `tolower()` to transform letters into small.

```{r}
# vector with pioneers and their birth year
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
# split them using strsplit() and inspect how it looks like before proceeding
split_data <- strsplit(pioneers, ":")
split_data

# now, turn all data into small letters by using lapply() and tolower():
split_low <- lapply(split_data, tolower)
# now reprint the data with small letters
split_low
```

By looking at the latest data, the letters were indeed transformed into small letters.

Do notice that the data returned by `lapply()` is a list.  To transform a list into an array, we can utilize `unlist()` to our finalized list called `split_low`

```{r}
# split_low, which we solved for above. is a list.
str(split_low)

# now unlist our data
unlist(split_low)
```

Now, all our data is in a nice sequential vector.  `unlist()` as stated is a great tool to simply data.

#### Using anonymous and custom functions in apply

Instead of using functions already in the workspace, we can pass an anonymous or our custom function to the family of apply functions.  Suppose we want to split the name and years of our data into two seperate datasets. Here is an example with using custom function I made from scratch and passing it as an anonymous an example:

```{r}
# Code from previous exercise:
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)

# Write function select_first()
select_first <- function(x) {
  x[1]
}
# PASSING A CUSTOM FUNCTION:  Applies select_first function over split_low and names it 'names'
names <- lapply(split_low, select_first)
names

# PASSING AN ANONYMOUS FUNCTION: Feeds split_low and an anonymous example to lapply.. call the result 'years'
years <- lapply(split_low, select_second)
years
```

#### sapply() method

`lapply()` returns a list that we can transform into a vector by using `unlist()`

But `sapply()` attempts to simplify our data into a vector or matrix.  It is supposed to be user-friendly.

Now let's play around some data and `sapply()`.  Suppose we have a vector of city names and we want to count the number of characters using the built-in `nchar()` method in R.

```{r}
cities <- c("London", "Prague", "Los Angeles", "Berlin")

# using lapply() to get a vector of character number
unlist(lapply(cities, nchar))

# using sapply() to get a vector of character number
sapply(cities, nchar)
```

The beauty of sapply is that the vector returned corresponds to the original data so we know which data corresponds to which data.

But if we do not want the names, we can pass USE.NAMES = FALSE as a paramater to `sapply()`.  Here is the example:

```{r}
sapply(cities, nchar, USE.NAMES = FALSE) 
```

#### Problem with sapply() when the function returns uneven vectors

The main issue with sapply() is that it forces to simply the results and the behavior can be unpredictable.  In the previous example, it simply returned a numerical value for each data.  But when you are processing multiple data output, the behavior can be predictable.  

Suppose we want to use `sapply()` to break down individual cities into their characters by using strsplit()
```{r}
cities <- c("London", "Prague", "Los Angeles", "Berlin")
# A function I created to split city names into individual characters
splitted_cities <- sapply(cities, function (city) {
  return(strsplit(city, ""))
})

# Returns the product of sapply()
splitted_cities 

# Prints the structure of this data
str(splitted_cities)

```

As revealed by the `str()` script, splitted_cities returned a list.  As stated earlier, `sapply()` forces the results into a vector or matrix.  But the unpredictability can sometimes force the result into a list if it does not know what to do. 

#### Detour: min() and max()

As a detour, we can use `min()` and `max()` to determine the lowest and highest dataset of a vector and matrix. 


#### Example:  sapply() when function returns vector of same length

When the anonymous function returns for instance the same length of vector, `sapply()` returns a beautiful labeled data matrix.  

Suppose we go back to the cities vector and want to determine the first and last letter of each city:  

```{r}
# Vector of city
cities <- c("London", "Prague", "Los Angeles", "Berlin")

# Applies an anonymous function via sapply() that determines the first and last letter of city.
city_sorter <- sapply (cities, function (city) {
  splitted_city <- strsplit(city, "")[[1]]
  c(first_letter = splitted_city[1], 
        last_letter = splitted_city[length(splitted_city)])
})
city_sorter
```

#### Detour:  cat()

`cat()` allows to concatenate variables, values together without the use of paste:

```{r}
weekend_profits <- list(January = c(122,218,213,500), February = c(1029,19212, 122))
message <- lapply(weekend_profits, function(month){
  cat("The mean profit for the month is", mean(month), "\n")
})
message
```



Notice that `cat()` printed the message we want, but it also returned unwanted data such as `$January` and `$February` with NULL values.  Not what we want!  Here comes `vapply()` to sidestep this problem.  

#### vapply()

`sapply()` has a number of problems as it forces to simplify the resulting data into a vector.  

On the other hand and to prevent this problem, `vapply()` requires to explictly specify the output we expect. Here is an example where numeric(1) is passed an argument to `vapply()` to specify the expected return value from the passed function.  

```{r}
# A vector of cities
cities <- c("London", "Prague", "Los Angeles", "Berlin")

# vapply() with specified numeric(1) input telling vapply that we expect a numeric value for each data
vapply(cities, nchar, numeric(1))
```

Another example from earlier using sapply() is the return of the first and last letter.  Let us go over the same task with vapply():

```{r}
cities <- c("London", "Prague", "Los Angeles", "Berlin")

# we modified the earlier sapply function with the added character(2) argument
city_sorter <- vapply (cities, function (city) {
  splitted_city <- strsplit(city, "")[[1]]
  c(first_letter = splitted_city[1], 
        last_letter = splitted_city[length(splitted_city)])
}, character(2))

city_sorter
```

Now, lets try to tell vapply that we're expecting only 1 character type as a return value even though the function returns two.

```{r}
# modified from above, now called city_sorter_2 with incorrect expected return value from callback function
city_sorter_2 <- vapply (cities, function (city) {
  splitted_city <- strsplit(city, "")[[1]]
  c(first_letter = splitted_city[1], 
        last_letter = splitted_city[length(splitted_city)])
}, character(1))
```

`vapply()` is prevented from completing the task and throws an error because it received values that deviated from what was expected.  Very conservative in a good way.  

#### apply family and passing another default value

Instead of setting an internal default value in the callback function, we can set it via the apply family function.  Use `vapply()` since it is considered the best among the three discussed for predictability and specificity. 

Here's an example where we increase the current price 

```{r}
# old price vector
price_vector <- c(eggroll = 2.50, boba = 1.50, popcorn_chicken = 5)

# new price vector with 10% price increase and tax inclusion
new_price_with_tax <- vapply(price_vector, function (price, tax) {
  round(price * 1.10 * (1 + tax), 2)
}, tax = 0.0625, numeric(1))

# prints the updated price with 10% increase and the tax rate as an argument to the vapply()
new_price_with_tax
```

