---
title: "Apply Family"
output: html_notebook
---

#### Basics of Apply

Apply allows us to simplify tasks by applying all items in a list/vector/matrix to a function without having to write a complicated for loop statement to iterate over the data structure we want to analyze.

#### Basics on apply

There's a family of apply function but let us examine `lapply()` and using it to split four data by cutting the semi-colon and making all entries small letters.

Again, we need to use `strsplit()` to split our character data and eventually need to use `tolower()` to transform letters into small.

```{r}
# vector with pioneers and their birth year
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
# split them using strsplit() and inspect how it looks like before proceeding
split_data <- strsplit(pioneers, ":")
split_data

# now, turn all data into small letters by using lapply() and tolower():
split_low <- lapply(split_data, tolower)
# now reprint the data with small letters
split_low
```

By looking at the latest data, the letters were indeed transformed into small letters.

Do notice that the data returned by `lapply()` is a list.  To transform a list into an array, we can utilize `unlist()` to our finalized list called `split_low`

```{r}
# split_low, which we solved for above. is a list.
str(split_low)

# now unlist our data
unlist(split_low)
```

Now, all our data is in a nice sequential vector.  `unlist()` as stated is a great tool to simply data.

#### Using anonymous and custom functions in apply

Instead of using functions already in the workspace, we can pass an anonymous or our custom function to the family of apply functions.  Suppose we want to split the name and years of our data into two seperate vectors. Here is an example with using custom functions I made from scratch:

```{r}
# Code from previous exercise:
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)
# Write function select_first()
select_first <- function(x) {
  x[1]
}
# Apply select_first() over split_low: names
names <- lapply(split_low, select_first)
# Write function select_second()
select_second <- function(x) {
  x[2]
}
# Apply select_second() over split_low: years
years <- lapply(split_low, select_second)

names

years
```




