---
title: "Tidyverse-Data Wrangling"
output:
  html_notebook: default
  html_document: default
---

#### Loading what we will need:

Our data is gapminder here.  And to use the piping notation (`%>%`) properly, we will need to use the dplyr library.  The piping notation should be read as THEN.  

```{r, message=FALSE, warning=FALSE}
# load gapminder package which is built within the R program
library(gapminder)
# load dplyr which is a key part of the tidyverse system
library(dplyr)

# call gapminder to see what kind of dataset it includes
gapminder
```

#### Filter verb 

Filter allows us to limit the results based on the parameters wanted which are passed on the `filter()` line of the code.  

__The task right now:__  Add a filter() line after the pipe (%>%) to extract only the observations from the year 1957. 

```{r}
# unnecessary since it was declared already but just put it for future use
library(gapminder)
library(dplyr)

# Filter the gapminder dataset for the year 1957
gapminder %>%
  filter(year == 1957)

# Filter for China in 2002
gapminder %>%
  filter(country == "China", year == 2002)
```

#### Arrange verb

It is used to sort data by its order:  either alphabetically or numerically.  We need to pass `arrange()` with what variable to sort by. The default of order is increasing.  If you need to override the default, passed variable must be wrapped by `desc()`.  Here's an example:

```{r}
# Sort in ascending order of lifeExp
gapminder %>% 
  arrange(lifeExp)

# Sort in descending order of lifeExp
gapminder %>% 
  arrange(desc(lifeExp))
```

#### Combining multiple verbs

Goal:  Use filter() to extract observations from just the year 1957, then use arrange() to sort in descending order of population (pop):

```{r}
gapminder %>%
  filter(year == 1957) %>%
  arrange(desc(pop))
```

#### Mutate verb

The mutate verb can be used to create a new column or mutate current columns.

__First task:__  Use `mutate()` to change the existing lifeExp column, by multiplying it by 12: 12 * lifeExp.

```{r}
gapminder %>%
  mutate(lifeExp = lifeExp * 12)
```

__Second task:__ Use `mutate()` to add a new column, called lifeExpMonths, calculated as 12 * lifeExp


```{r}
gapminder %>%
  mutate(lifeExpMonths = 12 * lifeExp)
```

#### Practice with combining all 3 verbs here:

In one sequence of pipes on the gapminder dataset:

- filter() for observations from the year 2007,
- mutate() to create a column lifeExpMonths, calculated as 12 * lifeExp, and
- arrange() in descending order of that new column

Here's the solution:


```{r}
# Filter, mutate, and arrange the gapminder dataset
gapminder %>%
  filter(year == 2007) %>%
  mutate(lifeExpMonths = 12 * lifeExp) %>%
  arrange(desc(lifeExpMonths))
```


