---
title: "Loops Notebook"
output: html_notebook
---

__Notes:__ Practically similar to Javascript and most OOP languages.

#### & and | 

Instead of double && and ||, we use the single version of them. Here's an example R script that utilizes these tools for relational operators.  

```{r}
# The linkedin and last variable are already defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
last <- tail(linkedin, 1)

# Is last under 5 or above 10?
last < 5 | last > 10

# Is last between 15 (exclusive) and 20 (inclusive)?
last > 15 & last <= 20
```

```{r}

```


#### Using truthy/falsey idea to count TRUE

Remember, TRUE counts as 1 and FALSE counts as 0.  We can use this to our advantage.

```{r}
# An array of TRUE and FALSE
logical_vector <- c(TRUE, FALSE, TRUE, FALSE, TRUE, TRUE)

# Use sum() to count TRUEs as a short, could write and if statement too as an alternative
sum(logical_vector)
```

#### if statement

Explanatory to other OOP langauges.  The following code is a basic script for an if statement with else if and else components:

```{r}
# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Control structure to check medium assuming we don't know it
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
} else if (medium == "Facebook") {
  # Add code to print correct string when condition is TRUE
  print("Showing Facebook information")
} else {
  print("Unknown medium")
}

# Control structure for num_views assuming we don't know it
if (num_views > 15) {
  print("You're popular!")
} else if (num_views <= 15 & num_views > 10) {
  # Add code to print correct string when condition is TRUE
  print("Your number of views is average")
} else {
  print("Try to be more visible!")
}
```


