---
title: "Loops Notebook"
output: html_notebook
---

__Notes:__ Practically similar to Javascript and most OOP languages.

#### & and | 

Instead of double && and ||, we use the single version of them. Here's an example R script that utilizes these tools for relational operators.  

```{r}
# The linkedin and last variable are already defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
last <- tail(linkedin, 1)

# Is last under 5 or above 10?
last < 5 | last > 10

# Is last between 15 (exclusive) and 20 (inclusive)?
last > 15 & last <= 20
```

```{r}

```


#### Using truthy/falsey idea to count TRUE

Remember, TRUE counts as 1 and FALSE counts as 0.  We can use this to our advantage.

```{r}
# An array of TRUE and FALSE
logical_vector <- c(TRUE, FALSE, TRUE, FALSE, TRUE, TRUE)

# Use sum() to count TRUEs as a short, could write and if statement too as an alternative
sum(logical_vector)
```

#### if statement

Explanatory to other OOP langauges.  The following code is a basic script for an if statement with else if and else components:

```{r}
# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Control structure to check medium assuming we don't know it
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
} else if (medium == "Facebook") {
  # Add code to print correct string when condition is TRUE
  print("Showing Facebook information")
} else {
  print("Unknown medium")
}

# Control structure for num_views assuming we don't know it
if (num_views > 15) {
  print("You're popular!")
} else if (num_views <= 15 & num_views > 10) {
  # Add code to print correct string when condition is TRUE
  print("Your number of views is average")
} else {
  print("Try to be more visible!")
}
```

#### while loop

Explanatory while in other OOP language.  Here's a short script example of a while loop in R:

```{r}
# Initialize the speed variable
speed <- 64
# Code the while loop
while (speed > 30) {
  print("Slow down!")
  speed <- speed - 7
}
# Print out the speed variable
speed
```

#### String concatenation using paste()

R lacks the basic ability to do string concatenation with the plus-sign like in most OOP languages.  So has to use the `paste()` method.  Here's a basic example of `paste()`:

```{r}
speed <- 64
paste("Your speed is", speed)
```


#### break statement 

Break statements allows us to abandon the control loop completely.  

Here's an example:

```{r}
# Initialize the speed variable
speed <- 88

while (speed > 30) {
  print(paste("Your speed is", speed))
  # Break the while loop when speed exceeds 80
  if (speed > 80) {
    break
  }
  # NOTE: Because of break, this wont get checked and while() is abandoned
  if (speed > 48) {
    print("Slow down big time!")
    speed <- speed - 11
  } else {
    print("Slow down!")
    speed <- speed - 6
  }
}
```

#### for loop

There's two ways to loop through a vector.

For both, let's add the numbers inside the vector without using the `sum()` method.

_Way 1:_  for (item in items) 

```{r}
nums_vector <- c(2, 3, 5, 7, 11, 13)
sum <- 0
# This adds each item to sum using
for (number in nums_vector) {
  sum <- sum + number
}
sum

```

_Way 2:_  for (i in 1:length(items))

```{r}
nums_vector <- c(2, 3, 5, 7, 11, 13)
sum <- 0
# Adding using for (i in 1:length(items))
for (i in 1:length(nums_vector)) {
  sum <- sum + nums_vector[i]
}
sum
```

Voila, both method yielded 41!  Both works perfectly as intended.


