---
title: "R Utilities"
output: html_notebook
---

The main purpose of this particular notebook is go over some basic utilities that I will use extensively in R.  

#### Mathematical Utilies

Some of the useful utilies are:

- `abs()`: Calculate the absolute value.
- `sum()`: Calculate the sum of all the values in a data structure.
- `mean()`: Calculate the arithmetic mean.
- `round()`: Round the values to 0 decimal places by default. 

#### Reversing a list

- `rev()` might be useful for this purpose of reversing a list.  Here is an example:

```{r}
number_vector <- c(1,9,5,0)
# reverses the order of any given vector
rev(number_vector)
```

#### Data Utilities 

Here are some useful data utilities in R:

- `seq()`: Generate sequences, by specifying the from, to, and by arguments.
- `rep()`: Repeat elements of vectors and lists.
- `sort()`: Sort a vector in ascending order. Works on numerics, but also on character strings and logicals.
- `str()`: Display the structure of any R object.
- `append()`: Merge vectors or lists. It's similar to splice.
- is.*(): Check for the class of an R object.
- as.*(): Convert an R object from one class to another.
- `unlist()`: Flatten (possibly embedded) lists to produce a vector.

```{r}
# Print a sequence: from 1 up to 20 by increments of 2
# ?seq gives:  seq(from = 1, to = 1, by = ((to - from)/(length.out - 1)) ...)
seq(from = 0, to = 20, by = 2) 

# Repeat a list twice (a,b,a,b) with each appending at the end
# ?rep shows:  rep(x, times = 1, length.out = NA, each = 1)
rep(c(1,2,3), times = 3)

# Repeat a list twice but copy is next to original
rep(c("a", "b"), each = 2)
```

