---
title: "R Utilities"
output: html_notebook
---

The main purpose of this particular notebook is go over some basic utilities that I will use extensively in R.  

#### Mathematical Utilies

Some of the useful utilies are:

- `abs()`: Calculate the absolute value.
- `sum()`: Calculate the sum of all the values in a data structure.
- `mean()`: Calculate the arithmetic mean.
- `round()`: Round the values to 0 decimal places by default. 

#### Reversing a list

- `rev()` might be useful for this purpose of reversing a list.  Here is an example:

```{r}
number_vector <- c(1,9,5,0)
# reverses the order of any given vector
rev(number_vector)
```

#### Data Utilities 

Here are some useful data utilities in R:

- `seq()`: Generate sequences, by specifying the from, to, and by arguments.
- `rep()`: Repeat elements of vectors and lists.
- `sort()`: Sort a vector in ascending order. Works on numerics, but also on character strings and logicals.
- `str()`: Display the structure of any R object.
- `append()`: Merge vectors or lists. It's similar to splice.
- is.*(): Check for the class of an R object.
- as.*(): Convert an R object from one class to another.
- `unlist()`: Flatten (possibly embedded) lists to produce a vector.

```{r}
# Print a sequence: from 1 up to 20 by increments of 2
# ?seq gives:  seq(from = 1, to = 1, by = ((to - from)/(length.out - 1)) ...)
seq(from = 0, to = 20, by = 2) 

# Print a sequence from 10 to 5 by increments of -0.5
seq(10, 5, -0.5)

# Repeat a list twice (a,b,a,b) with each appending at the end
# ?rep shows:  rep(x, times = 1, length.out = NA, each = 1)
rep(c(1,2,3), times = 3)

# Repeat a list twice but copy is next to original
rep(c("a", "b"), each = 2)

# Combine 2 list in index 3 of the first list
# str(append) gives: function (x, values, after = length(x)) 
append(c(1,2,3,4,5), c("a", "b"), after = 2)
```


#### Note using sum() and the plus sign (+)

Suppose we have two vectors (1,1,1,1,1) and (1,2) and want to get the sum of all the vector entries. the order of sum and + matters.  

```{r}
# Two vectors of interest.  The sum of all items is 8 by inspection
vec1 <- c(1,1,1,1,1)
vec2 <- c(1,2)

# The incorrect way and R will give an answer with an error.  dim vec 1 =/= dim vec2 
sum(vec1 + vec2) 

# Two correct ways which will both yield 8
sum(c(vec1, vec2))
sum(vec1) + sum(vec2)
```

#### Regex using grepl() and grep()

`grepl()` returns a logical--makes sense given the l
whereas `grep()` gives a vector of indices where the pattern is found in a vector which is useful for subsetting.  If the variable given to grep() is of length 1, it will just return 1.  

Before I do tasks with `grepl()`, I will examine the str of grepl to see the parameters involved:

```{r}
str(grepl)
function (pattern, x, ignore.case = FALSE, perl = FALSE, fixed = FALSE, 
    useBytes = FALSE)
```

Knowing the basic structure, let's go over this basic code and what each do:

```{r}
# An emails vector
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "dalai.lama@peace.org",
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")

# Use grepl() to match for "edu", it returns a logical vector
grepl("edu", emails)

# Use grep() to match for "edu", saves result in vector with matched indeces
hits <- grep("edu", emails)

# Subset emails using hits
emails[hits]

# Note:  Subsetting can be fed with a vector of indeces like above 
# But subsetting also works with a vector of logical values
```

#### Regex - better filter for previous results

Suppose we want to filter the data more accurately:

We want to filter data via:  "\@name.edu":

```{r}
# vector named emails was already declared in the preceding R script
# Use grepl() to match for the more specific filter
grepl("@.*\\.edu$", emails)
```

Explanation of the filter is as follows:
- `@` is self explanatory
- `.*` has two parts: `.` is the character class which looks for any non-next line character, and `*` is the quantifier which implies to search for any length.  so `.*` is valid for searching any domain name between the `@` and `.` sign
- `\\.` is the equivalent of a simple dot in R's regex
- `$` signifies the end of the string.  so the result has to end with this.


#### Substitution through sub() and gsub()

`sub()` and `gsub()` are powerful tools.  `sub()` only replaces the first match in an index, whereas `gsub()` will replace every single item. Note that the replacement string does not require to be in regex format.

Here is an example where we replace the ending of any `@__.edu` to `@test.edu`:

```{r}
# The emails vector 
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "global@peace.org",
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")
# Using sub ... again, replacement argument should not be in regex format
sub("@.*\\.edu$", "@test.edu", emails)
```


#### sub using parenthesis format

Suppose we want to replace values that includes AA .. Az with the second letter only.  The appropriate regex format is:  "A([A-z])", which requires the parenthesis. Parenthesis is remembered which can be invoked as "\\1" in the replacement part.  

```{r}
# Replace parts with AA .. Az with the second letter only (A...z)
gsub("A([A-z])", "\\1", "12AB99Az")
```

Another example would be to extract the number of nominations from a vector called awards.  Each index contains the ending "... nominations".  Here is a sample code:

```{r}
# a vector called awards with dummy data on number of nominations
awards <- c("Won 1 Oscar. Another 9 wins & 24 nominations.",
  "1 win and 2 nominations.",
  "2 wins & 3 nominations.",
  "Nominated for 2 Golden Globes. 1 more win & 2 nominations.",
  "4 wins & 1 nomination.")

nominations <- sub(".*\\s([0-9]+)\\snomination.*$", "\\1", awards)
# Should print the number of nominations each index has.  
nominations
```


#### Sys.___  to get time date

Sys.Date() can be used to get current data, Sys.time() for the time

Sys.Date() is an instance of Date object.  Sys.time() is an instance of POSIXct object.

```{r}
# Get the current date: today
today <- Sys.Date()
today

# Get the current time: now
now <-  Sys.time()
now
```


#### Data formatting in R:

The character string has to obey a format that can be defined using a set of symbols (the examples correspond to 13 January, 1982):


- %Y: 4-digit year (1982)
- %y: 2-digit year (82)
- %m: 2-digit month (01)
- %d: 2-digit day of the month (13)
- %A: weekday (Wednesday)
- %a: abbreviated weekday (Wed)
- %B: month (January)
- %b: abbreviated month (Jan)

NOTE: The main advantage of putting it in unix timestamp (POSIXct) is that we do basic mathematical operation with them (adding date, determining differences)

#### Forcing stored date through as.Date()

Remember, date is stored in the "%Y-%m-%d" (e.g. "2018-07-04").  If string happens to be in this format, no need to specify format.  Otherwise, a specific format using the appropriate '%' is required to be passed to `as.Date()`.  Moreover, once a data object, we can simply use `format()` to print it a certain way.  

```{r}
# 3 different date formats we have as string and converting them to unix timestamp
date_char1 <- "1982-01-13"
as.Date(date_char1)

date_char2 <- "Jan-13-82"
as.Date(date_char2, format = "%b-%d-%y")

date_char3 <- "13 January, 1982"
to_format <- as.Date(date_char3, format = "%d %B, %Y")

# Suppose we want to print to_format, which is now Date object as "MM YY" format
format(to_format, "%m %y")
```



