---
title: "Functions in R"
output: html_notebook
---

#### Looking up documentation for built-in functions/methods

? is useful.  But an equivalent way to search is by using `help()`

In R, ?sd and help(sd) are the same to look up thorough documentation on SD.  

#### Shortcut using args() 

Suppose we want to see the arguments and default positions on a method is.  We can use `args()` -- we can technically use str() too!

```{r}
# Using args(), this looks up the variables we can pass, their inherent order, and their default values for data.frame() which we use to create a df object 
args(data.frame)


# Using str(), this looks up the variables we can pass, their inherent order, and their default values for matrix() which we use to create a matrix object
str(matrix)
```

#### mean() and removing NA values

If you look at the documentation of mean, it gives this as a structure:

mean(x, trim = 0, na.rm = FALSE, ...)

Suppose we want to calculate the mean for a vector that includes a NA value, `mean()` would then return a NA value which is not what we want.  `na.rm = FALSE` is the default but passing `na.rm = TRUE` to the `mean()` method would ignore the NA value. Here's the script showing this paragraph in code:

```{r}
linkedin <- c(16, 9, 13, 5, NA, 17, 14)

# default for mean, does not work
mean(linkedin)

# passing na.rm = TRUE to ignore NA values, does work
mean(linkedin, na.rm = TRUE)
```

#### Writing own function

R again has much in common with programs such as JS.  Here is a basic function in R:  

```{r}
# Create a function pow_two() that multiplies that arg to itself
pow_two <- function (arg) { 
  arg * arg
}
# Use the function
pow_two(12)

# Create a function sum_abs() that adds the abs val of each argument
sum_abs <- function (arg1, arg2) {
  abs(arg1) + abs(arg2)
}
# Use the function
sum_abs(-2,3)
```

#### Difference between returning and printing:

Before I explain, let's examine this code without explanation:

```{r}
# Define the function hello()
hello <- function () {
  print("Hi there!")
  TRUE 
  # Note:  return() can be used to be more explicit but leaving TRUE works.  return with parenthesis unlike in JS
}

what_is_returned <- hello()
what_is_returned
```


The code above prints the hi there message but it returns a TRUE function when it is invoked with hello().

#### Writing optional argument with default

Assignment statement in R utilizes the <- symbol.  But assigning a default argument utilizes the classic = symbol as assignment.  


```{r}
# print_info is TRUE by default
pow_two <- function(x, print_info = TRUE) {
  y <- x ^ 2
  if (print_info == TRUE) {
    print(paste(x, "to the power two equals", y))
  }
  #Again return has to wrap the argument with parenthesis
  return(y) 
}

# invoking the function with one argument
pow_two(2)

# invoking the function with two arguments (overriding the default)
pow_two(3, FALSE)
```

#### Packages

Methods we've been using come from installed packages.  Base package for instance are installed.  But packages like ggivis are not pre-installed. So have to install manually.

#### search() to see which package is installed

Running `search()` in the console will reveal all the packages installed in our module.

```{r}
search()
```

#### Installing packages that are not installed

Before we can utilize a new package, we need to install that package and then load it to our current session.  Suppose we want to install ggvis:

```{r, message=FALSE, warning=FALSE}
# Install R 
install.packages("ggvis")
```

The ggvis package we installed is still not available in our R session--`search()` will not show ggvis in the list.  For that we need to load it using `library()`

```{r}
#loading ggvis into our R session (either with quote or no quote)
library(ggvis)

#Using search() to verify that ggvis is now available
search()
```

Indeed, `search()` now shows package:ggvis being available to our current R session.

#### library() vs require()

`require()` and `library()` behave the same.  The only difference is that `require()` will return a FALSE logical statement if the loading fails.  

#### Ways to install review

```
# Chunk 1: Valid, works without quotes
library(data.table)
require(rjson)

# Chunk 2:  Valid, works with quotes
library("data.table")
require(rjson)

# Chunk 3:  Not valid, might seem valid but passing another argument to require forces us to surround rjson with quotes
library(data.table)
require(rjson, character.only = TRUE)

# Chunk 4:  Not valid once library() doc is examined for valid input
library(c("data.table", "rjson"))
```

