---
title: "Functions in R"
output: html_notebook
---

#### Looking up documentation for built-in functions/methods

? is useful.  But an equivalent way to search is by using `help()`

In R, ?sd and help(sd) are the same to look up thorough documentation on SD.  

#### Shortcut using args() 

Suppose we want to see the arguments and default positions on a method is.  We can use `args()` -- we can technically use str() too!

```{r}
# Using args(), this looks up the variables we can pass, their inherent order, and their default values for data.frame() which we use to create a df object 
args(data.frame)


# Using str(), this looks up the variables we can pass, their inherent order, and their default values for matrix() which we use to create a matrix object
str(matrix)
```

#### mean() and removing NA values

If you look at the documentation of mean, it gives this as a structure:

mean(x, trim = 0, na.rm = FALSE, ...)

Suppose we want to calculate the mean for a vector that includes a NA value, `mean()` would then return a NA value which is not what we want.  `na.rm = FALSE` is the default but passing `na.rm = TRUE` to the `mean()` method would ignore the NA value. Here's the script showing this paragraph in code:

```{r}
linkedin <- c(16, 9, 13, 5, NA, 17, 14)

# default for mean, does not work
mean(linkedin)

# passing na.rm = TRUE to ignore NA values, does work
mean(linkedin, na.rm = TRUE)
```

