---
title: "Intro to Basics"
output: html_notebook
---



### Some basic arithmetic in R


```{r}
# An addition
5 + 5 

# A subtraction
5 - 5 

# A multiplication
3 * 5

 # A division
(5 + 5) / 2 

# Exponentiation
2 ^ 5

# Modulo
28 %% 6
```

__Self Note:__  Exponentiation is similar with Javascript, but different from Python (which uses **).  Modulo on the other hand is unique since it is double %%.  



### Variable Assignment


The script below assigns four to variable called my_var:
```{r}
my_var <- 4
my_var
```

#### Adding variables
```{r}
# Assign a value to the variables my_apples and my_oranges
my_apples <- 5
my_oranges <- 6 


# Add these two variables together
my_apples + my_oranges

# Create the variable my_fruit
my_fruit <- my_apples + my_oranges
```

#### Data Type

- Decimals values like 4.5 are called __numerics__.
- Natural numbers like 4 are called integers. Integers - are also __numerics__.
- Boolean values (TRUE or FALSE) are called __logical__.
- Text (or string) values are called __characters__.

```{r}
# Declare variables of different types
my_numeric <- 42
my_character <- "universe"
my_logical <- FALSE 

# Check class of my_numeric
class(my_numeric)

# Check class of my_character
class(my_character)

# Check class of my_logical
class(my_logical)
```

__Self Note:__  class of strings is "character", class of numbers is "numeric", and class of boolean is "logical"

#### Vectors

`c()` is the method to create a vector. The c in `c()` stands for combine.  

```{r}

```
