---
title: "Intro to Basics"
output: html_notebook
---



### Some basic arithmetic in R


```{r}
# An addition
5 + 5 

# A subtraction
5 - 5 

# A multiplication
3 * 5

 # A division
(5 + 5) / 2 

# Exponentiation
2 ^ 5

# Modulo
28 %% 6
```

\ 

\

__Self Note:__  Exponentiation is similar with Javascript, but different from Python (which uses **).  Modulo on the other hand is unique since it is double %%.  



### Variable Assignment


The script below assigns four to variable called my_var:
```{r}
my_var <- 4
my_var
```

#### Adding variables
```{r}
# Assign a value to the variables my_apples and my_oranges
my_apples <- 5
my_oranges <- 6 


# Add these two variables together
my_apples + my_oranges

# Create the variable my_fruit
my_fruit <- my_apples + my_oranges
```

#### Data Type

- Decimals values like 4.5 are called numerics.
- Natural numbers like 4 are called integers. Integers - are also numerics.
- Boolean values (TRUE or FALSE) are called logical.
- Text (or string) values are called characters.
