---
title: "Intro to Basics"
output: html_notebook
---



### Some basic arithmetic in R


```{r}
# An addition
5 + 5 

# A subtraction
5 - 5 

# A multiplication
3 * 5

 # A division
(5 + 5) / 2 

# Exponentiation
2 ^ 5

# Modulo
28 %% 6
```

__Self Note:__  Exponentiation is similar with Javascript, but different from Python (which uses **).  Modulo on the other hand is unique since it is double %%.  



### Variable Assignment


The script below assigns four to variable called my_var:
```{r}
my_var <- 4
my_var
```

#### Adding variables
```{r}
# Assign a value to the variables my_apples and my_oranges
my_apples <- 5
my_oranges <- 6 


# Add these two variables together
my_apples + my_oranges

# Create the variable my_fruit
my_fruit <- my_apples + my_oranges
```

#### Data Type

- Decimals values like 4.5 are called __numerics__.
- Natural numbers like 4 are called integers. Integers - are also __numerics__.
- Boolean values (TRUE or FALSE) are called __logical__.
- Text (or string) values are called __characters__.

```{r}
# Declare variables of different types
my_numeric <- 42
my_character <- "universe"
my_logical <- FALSE 

# Check class of my_numeric
class(my_numeric)

# Check class of my_character
class(my_character)

# Check class of my_logical
class(my_logical)
```

__Self Note:__  class of strings is "character", class of numbers is "numeric", and class of boolean is "logical"

#### Vectors

`c()` is the method to create a vector. The c in `c()` stands for combine.  Vectors can hold numeric, character, and logical.  Example:

```{r}
numeric_vector <- c(1, 10, 49)
character_vector <- c("a", "b", "c")
boolean_vector <- c(TRUE, FALSE, TRUE)
```

#### Naming pre-existing vectors

Naming vectors are important to keep track of what the data you're holding is for.  For instance, a vector could hold a series of data (e.g. earnings) for each day.  But we need a way to name which data corresponds to which. The code below illustrates the labeling standard.

```{r}
 # Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)
# Assign days as names of poker_vector
names(poker_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
# Assign days as names of roulette_vectors
names(roulette_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
```

