---
title: "Import Data--Flat Files"
output: html_notebook
---

#### What is a flat file?

Flat file like .csv is a file with data separated by a `,`. 

Can use `utils::read.csv()` to read file

#### read.csv()

- Sample way to read csv file is:
  `pools <- read.csv("swimming_pools.csv")`
- To see the basic structure of imported data, use `str(pools)` to see levels, variables, etc

```{r}
pools <- read.csv("swimming_pools.csv")
```

#### read.csv(), forcing it to NOT convert variables into categorical or factors

Note:  All importing functions in utils package, such as `read.csv()`, imported strings are converted to factors.  

The correct way would be to run the following code:
`pools<- read.csv("swimming_pools.csv", stringsAsFactors = FALSE)`

```{r}
proper_pools <- read.csv("swimming_pools.csv", stringsAsFactors = FALSE)
improper_pools <- read.csv("swimming_pools.csv")
```

To see the difference, compare the str() printout from two imported data:

The code below shows that the improper way of importing data without the `stringsAsFactors = FALSE` argument yields to name and address being converted to factors.  

```{r}
str(improper_pools)
```

The code below shows that name and address variables are kept as characters which is the desired and proper way of retrieving and saving data. 

```{r}
str(proper_pools)
```


#### read.delim and read.table

- `read.delim()` is used for tab-delimited
- `read.table()` is used for any flat file

Tasks: 
- Import the data in "hotdogs.txt" with read.delim(). Call the resulting data frame hotdogs. The variable names are not on the first line, so make sure to set the header argument appropriately.
- Call summary() on hotdogs. This will print out some summary statistics about all variables in the data frame.

__Note__:  header = FALSE ensures that the first line is not treated as variable names.

```{r}
hotdogs <- read.delim("hotdogs.txt", header = FALSE)
summary(hotdogs)
```

#### Using read.table()

An example of using `read.table()` with tab-seperated data:

```{r}
# Path to the hotdogs.txt file: path
path <- file.path("data", "hotdogs.txt")

# Import the hotdogs.txt file: hotdogs
hotdogs <- read.table(path,  
                      sep ="\t",
                      col.names = c("type", "calories", "sodium"))

# Call head() on hotdogs
head(hotdogs)
```


#### Exercise:  arguments on calories vs sodium

__Tasks:__ Finish the read.delim() call to import the data in "hotdogs.txt". It's a tab-delimited file without names in the first row. The code that selects the observation with the lowest calorie count and stores it in the variable lily is already available. It uses the function which.min(), that returns the index the smallest value in a vector.
Do a similar thing for Tom: select the observation with the most sodium and store it in tom. Use which.max() this time. Finally, print both the observations lily and tom.

```{r}
# Finish the read.delim() call
hotdogs <- read.delim("hotdogs.txt", header = FALSE, col.names = c("type", "calories", "sodium"))

# Select the hot dog with the least calories: lily
lily <- hotdogs[which.min(hotdogs$calories), ]

# Select the observation with the most sodium: tom
tom <- hotdogs[which.max(hotdogs$sodium), ]

# Print lily and tom
lily
tom
```

__Results:__ Based on the result above, Lily would get the 50th entry with Poulty.  Whereas Tom would get the 15th hotdog with Beef.


