---
title: "Import Data--Flat Files"
output: html_notebook
---

#### What is a flat file?

Flat file like .csv is a file with data separated by a `,`. 

Can use `utils::read.csv()` to read file

#### read.csv()

- Sample way to read csv file is:
  `pools <- read.csv("swimming_pools.csv")`
- To see the basic structure of imported data, use `str(pools)` to see levels, variables, etc

```{r}
pools <- read.csv("swimming_pools.csv")
```

#### read.csv(), forcing it to NOT convert variables into categorical or factors

Note:  All importing functions in utils package, such as `read.csv()`, imported strings are converted to factors.  

The correct way would be to run the following code:
`pools<- read.csv("swimming_pools.csv", stringsAsFactors = FALSE)`

```{r}
proper_pools <- read.csv("swimming_pools.csv", stringsAsFactors = FALSE)
improper_pools <- read.csv("swimming_pools.csv")
```

To see the difference, compare the str() printout from two imported data:

The code below shows that the improper way of importing data without the `stringsAsFactors = FALSE` argument yields to name and address being converted to factors.  

```{r}
str(improper_pools)
```

The code below shows that name and address variables are kept as characters which is the desired and proper way of retrieving and saving data. 

```{r}
str(proper_pools)
```


#### read.delim and read.table

- `read.delim()` is used for tab-delimited
- `read.table()` is used for any flat file




