---
title: "Data Visualization"
output: html_notebook
---

#### Visualization using ggplot2 library

Before we can plot, we need to load the ggplot2 library first:

```{r}
library(ggplot2)

# we will need to manipulate data so load dplyr.
library(dplyr)
# load gapminder data we previously used
library(gapminder)
```


#### Practice: scatter plot via ggplot() + geom_point():

Suppose we filter the data from gapminder for data in year 1952.  Using this filtered data, we want to create a plot with pop against gdpPercap:

```{r}
# Filtering data and assigning the results to a variable
gapminder_1952 <- gapminder %>%
  filter(year == 1952)

# Change to put pop on the x-axis and gdpPercap on the y-axis.  
# Note:  geom_point() is what plots it like: ggplot() + geom_point()
ggplot(gapminder_1952, aes(x = pop, y = gdpPercap)) +
  geom_point()
```

#### Change axis to log by adding:  scale_x_log10()

Using the previously declared and filtered data, we want to create a log-10 x scale for scatter plot with x as population (pop) and y as life expectancy (lifeExp):

```{r}
# we previously filtered and declared gapminder_1952
# reminder: regular plotting requires ggplot() + geom_point()

# plotting data with log10 as the primary scale for x-axis:
ggplot(gapminder_1952, aes(pop , lifeExp)) + geom_point() + scale_x_log10()
```

#### Color aesthetic

A simple way to add color to our previous graphs is by passing `color = continent` inside aes() as an argument.  Let's see what happens to our graph:

```{r}
ggplot(gapminder_1952, aes(pop , lifeExp, color = continent)) + geom_point() + scale_x_log10()
```



