---
title: "Matrix"
output: html_notebook
---

#### Constructing a matrix

Like vectors, all entries must be the same.  We can declare a vector by using `matrix()`

To construct a matrix with 1-9, which is filled by finishing up a row by row, we must use the following code:

```{r}
matrix(1:9, byrow= TRUE, nrow=3)
```

Note that the the default for `byrow` is FALSE.  Let's see what happens when we do not override the default:

```{r}
matrix(1:9, nrow=3)
```

#### Labeling matrix using colnames() and rownames()

To label matrices, we need to create a vector with the labels.  And then attach them to our matrix data using `colnames()` and `rownames()` as demonsrated below:

```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)
# Construct matrix
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), nrow = 3, byrow = TRUE)
# Vectors region and titles, used for naming
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# Name the columns with region
colnames(star_wars_matrix) <- region
# Name the rows with titles
rownames(star_wars_matrix) <- titles
# Print out star_wars_matrix
star_wars_matrix
```

#### Adding rows using `rowSums()`

Suppose we want to add both US and non-US numbers to get a sum.  Using the code from the previous page, we can simply use rowSums() to get that number:

```{r}
rowSums(star_wars_matrix)
```

#### Creating new matrix via `cbind()`

Suppose we want to create a new matrix using the results of appending the sums into our dataset, star_wars_matrix.  This requires adding another column to our matrix which we can do using `cbind()`

```{r}
sum_vector <- rowSums(star_wars_matrix)
all_wars_matrix <- cbind(star_wars_matrix, sum_vector)
all_wars_matrix 
```

Using `cbind()`, I created a new matrix called all_wars_matrix with the combined dataset.  

__Note:__  To add a new matrix by adding 2 or more datasets by row, `rbind()` can be utilized like we did with `cbind()`
