---
title: "Matrix"
output: html_notebook
---

#### Constructing a matrix

Like vectors, all entries must be the same.  We can declare a vector by using `matrix()`

To construct a matrix with 1-9, which is filled by finishing up a row by row, we must use the following code:

```{r}
matrix(1:9, byrow= TRUE, nrow=3)
```

Note that the the default for `byrow` is FALSE.  Let's see what happens when we do not override the default:

```{r}
matrix(1:9, nrow=3)
```

#### Labeling matrix using colnames() and rownames()

To label matrices, we need to create a vector with the labels.  And then attach them to our matrix data using `colnames()` and `rownames()` as demonsrated below:

```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)
# Construct matrix
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), nrow = 3, byrow = TRUE)
# Vectors region and titles, used for naming
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# Name the columns with region
colnames(star_wars_matrix) <- region
# Name the rows with titles
rownames(star_wars_matrix) <- titles
# Print out star_wars_matrix
star_wars_matrix
```

#### Adding rows using `rowSums()`

Suppose we want to add both US and non-US numbers to get a sum.  Using the code from the previous page, we can simply use rowSums() to get that number:

```{r}
rowSums(star_wars_matrix)
```

#### Creating new matrix via `cbind()`

Suppose we want to create a new matrix using the results of appending the sums into our dataset, star_wars_matrix.  This requires adding another column to our matrix which we can do using `cbind()`

```{r}
sum_vector <- rowSums(star_wars_matrix)
all_wars_matrix <- cbind(star_wars_matrix, sum_vector)
all_wars_matrix 
```

Using `cbind()`, I created a new matrix called all_wars_matrix with the combined dataset.  

__Note:__  To add a new matrix by adding 2 or more datasets by row, `rbind()` can be utilized like we did with `cbind()`

#### Selecting matrix elements

my_matrix[1,2] selects the element at the first row and second column.

my_matrix[1:3,2:4] results in a matrix with the data on the rows 1, 2, 3 and columns 2, 3, 4.

my_matrix[1,] selects all elements of the first row while my_matrix[,1] selects all elements of the first column.

Here's some practice with notes:

```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Select the non-US revenue for all movies
non_us_all <- all_wars_matrix[,2]
  
# Average non-US revenue
mean(non_us_all)

# Select the non-US revenue for first two movies
non_us_some <- non_us_all[1:2]
  
# Average non-US revenue for first two movies
mean(non_us_some)
```

#### Matrix arithmetic

For example, 2 * my_matrix multiplies each element of my_matrix by two. 
We can also multiply or divide another matrix with another matrix.  The matrix must be the same size.  And each component from each index of the matrix will perform the desired operation with another component from the same index from the other matrix.

