---
title: "Grouping and Summarizing Notebook"
output: html_notebook
---

#### Housekeeping:

Like in the previous sections for tideverse, we will use gapminder and dplyr:  
```{r}
library(gapminder)
library(dplyr)
```

#### Summarize verb

`summarize()` from tidyverse allows to print variable(s) at the end of piping operations.  Suppose we don't want to print all other data except the median of a variable in that dataset.  

Suppose we want to summarize the median of life expectancy which we will call medianLifeExp:

```{r}
gapminder %>%
  summarize(medianLifeExp = median(lifeExp))
```

#### Another practice by using other tidyverse verbs:

Task:  Find both the median life expectancy (lifeExp) and the maximum GDP per capita (gdpPercap) in the year 1957, calling them medianLifeExp and maxGdpPercap respectively. You can use the `max()` function to find the maximum.

```{r}
gapminder %>%
  filter(year == 1957) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))
```

#### Grouping via group_by()

Suppose we want to know the median life expectancy and max GDP for each year, then we can use `group_by()` before we invoke `summarize()`.  Here's an example:

```{r}
gapminder %>%
  group_by(year) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))
```

As shown above, `group_by()` allowed us to partition the summarize verb results for each year that exists in our dataset.

#### Using more than one filter for group_by()

We can use more than one filter for our group verb and it will find each unique combination of both.  

Suppose we want to ind the median life expectancy (lifeExp) and maximum GDP per capita (gdpPercap) within each combination of continent and year, saving them into medianLifeExp and maxGdpPercap, respectively:

```{r}
gapminder %>%
  group_by(continent, year) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))
```


