---
title: "Data Frame Section"
output: html_notebook
---

#### Data Frame Basics

Data Frame, unless matrix, can handle multiple data types.  

An example of data frame is `mtcars` that is built within R.  Calling `mtcars` invokes the following:


```{r}
mtcars
```

Random fact:  The class of data frame objects per `class()` method is "data.frame" as shown by the script below:

```{r}
class(mtcars)
```

#### Viewing the head and tail of data sets

If you are curious to see both the head and tail of data sets, `head()` and `tail()` are particularly useful methods for this purpose.


```{r}
head(mtcars)
```

Now for the bottom tail of the data set:

```{r}
tail(mtcars)
```

#### Utilizing structure method str() 

Suppose we want to see a summary of structures of data frames, `str()` is particularly useful here for insight.  `summary()` is not particularly good for this one as it will produce a bunch of extra non-indeed about structures.

Here's what `str()` would show you:

```{r}
str(mtcars)
```


The `str()` method demonstrates that there are 32 observations, 11 variables, and the corresponding data type for each variables--which happens to be numerical data type for all values here.  

#### Creating dataframes from scratch using data.frame()

To create a data frame, we can simply pass the columns we want to include into `data.frame()`.  The following is a rigorous example for a simple planetary data collection:

```{r}
# Definition of vectors
name <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)
# Create a data frame from the vectors
planets_df <- data.frame(name, type, diameter, rotation, rings)

# Now, let us invoke planets_df to see the constructed data frame
planets_df
```

#### Data Selection:

Like in matrices, we can easily select datasets or a range of them.  

Working on the previously declared planets_df dataframe we created, we can select and run the following to target wanted data:

```{r}
# Print out data for Mars (entire fourth row)
planets_df[4,]

# Print out diameter of Mercury (row 1, column 3)
planets_df[1,3]
```

#### Data Selection by using variable name df_name(,"var_name")

Suppose we want to search values using a known variable name, we can use the name as a filter to select a column:

```{r}
# Select first 5 values of diameter column
planets_df[1:5,"diameter"]
```


#### Data Selection by using variable name df_name(,c("var_name1", "var_name2"))

Another better way is to include names of the planets to see which of these values correspond to. For that, we can pass a vector of variable names. 

```{r}
# Select first 5 values of diameter column with their corresponding planets
planets_df[1:5,c("name","diameter")]
```

Voila, diameters were printed along with their associated planet names.


