---
title: "Data Frame Section"
output: html_notebook
---

#### Data Frame Basics

Data Frame, unless matrix, can handle multiple data types.  

An example of data frame is `mtcars` that is built within R.  Calling `mtcars` invokes the following:


```{r}
mtcars
```

Random fact:  The class of data frame objects per `class()` method is "data.frame" as shown by the script below:

```{r}
class(mtcars)
```

#### Viewing the head and tail of data sets

If you are curious to see both the head and tail of data sets, `head()` and `tail()` are particularly useful methods for this purpose.


```{r}
head(mtcars)
```

Now for the bottom tail of the data set:

```{r}
tail(mtcars)
```

#### Utilizing structure method str() 

Suppose we want to see a summary of structures of data frames, `str()` is particularly useful here for insight.  `summary()` is not particularly good for this one as it will produce a bunch of extra non-indeed about structures.

Here's what `str()` would show you:

```{r}
str(mtcars)
```


The `str()` method demonstrates that there are 32 observations, 11 variables, and the corresponding data type for each variables--which happens to be numerical data type for all values here.  

