---
title: "Visualization Types"
output: html_notebook
---

#### Create Line Plot via geom_line()

The syntax is pretty similar to the `geom_plot()` we've used earlier.  But instead, we use `geom_line()`

Task's subtask #1:  Use group_by() and summarize() to find the median GDP per capita within each year, calling the output column medianGdpPercap. Use the assignment operator <- to save it to a dataset called by_year.

Task's subtask #2: Use the by_year dataset to create a line plot showing the change in median GDP per capita over time. Be sure to use expand_limits(y = 0) to include 0 on the y-axis.

```{r}
# Summarize the median gdpPercap by year and continent , then save it as by_year
by_year <- gapminder %>%
  group_by(year, continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))
# Create a line plot showing the change in medianGdpPercap over time, and add color per continent
ggplot(by_year, aes(year, medianGdpPercap, color = continent)) + geom_line() + expand_limits(y=0)
```

#### Create Bar Plot via geom_col()

__Task 1:__ Use group_by() and summarize() to find the median GDP per capita within each continent in the year 1952, calling the output column medianGdpPercap. Use the assignment operator <- to save it to a dataset called by_continent.

__Task 2:__ Use the by_continent dataset to create a bar plot showing the median GDP per capita in each continent.

```{r}
by_continent <- gapminder %>%
  filter(year ==1952) %>%
  group_by(continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))

ggplot(by_continent, aes(continent, medianGdpPercap)) + geom_col() 
```