---
title: "Naive Bayes"
output: html_notebook
---

## __NOTE__:  data sets are not present here since they're proprietary data but includes my basic codes for tasks I accomplished

#### Basics and how to retrieve basic probabilities

```{r}
# Compute P(A) 
p_A <- nrow(subset(where9am, location == "office"))/nrow(where9am)

# Compute P(B)
p_B <- nrow(subset(where9am, daytype == "weekday"))/nrow(where9am)

# Compute the observed P(A and B)
p_AB <- nrow(subset(where9am, location == "office", daytype == "weekday"))/nrow(where9am)

# Compute P(A | B) and print its value
p_A_given_B <- p_AB/p_B
```


#### Working with Naive Bayes package

```{r}
# Load the naivebayes package
library(naivebayes)

# Build the location prediction model to predict location by using daytype independent var
locmodel <- naive_bayes(location ~ daytype, data = where9am)

# Predict Thursday's 9am location
predict(locmodel , thursday9am)
## will print: office

# Predict Saturdays's 9am location
predict(locmodel , saturday9am)
## will print: home
```

#### Type = "prob" for naive_bayes



```{r}
# Suppose the Naive Bayes 'locmodel' has been built

# Examine the location prediction model
locmodel
## Prints the following:
## daytype   appointment    campus      home    office
##   weekday   1.0000000 1.0000000 0.3658537 1.0000000
##   weekend   0.0000000 0.0000000 0.6341463 0.0000000

# Obtain the predicted probabilities for Thursday at 9am
predict(locmodel, thursday9am , type = "prob")
## Prints the following:
## appointment    campus      home office
##[1,]  0.01538462 0.1538462 0.2307692    0.6
```

__Conclusion:__ without the `type = "prob"` option, we will just get the most probable as output, not the probability distribution.

#### Why called 'naive'

A-B-C intersection = A-B intersection  * B-C intersection

But a descent approximation.  Assume independence so we can simply multiply and no worry about minusing intersection probabilities.

The 'infrequent' problem is when for instance we have no data on B-C so P(B-C) is 0 in this case.  So we can do a Laplace correction by adding 1 to each P(intersection)

#### Multiple independent variable

The dataset locations is already loaded in your workspace. Use the R formula interface to build a model where location depends on both daytype and hourtype. Recall that the function naive_bayes() takes 2 arguments: formula and data. Predict Brett's location on a weekday afternoon using the dataframe weekday_afternoon and the predict() function.

```{r}
# Build a NB model of location
locmodel <- naive_bayes(location ~ daytype + hourtype, data=locations)

# Predict Brett's location on a weekday afternoon
predict(locmodel, weekday_afternoon)
# side node:  weekday_afternoon is predefined as:
#   daytype  hourtype
#  weekday afternoon
## output: office

```







