---
title: "Logistic Regression"
output: html_notebook
---

#### Note:  Dataset is proprietary so don't have the dataset here, just the scripts I ran to solve the problems in another platform

#### Logistic Example

```{r}
# Examine the dataset to identify potential independent variables
str(donors)

# Explore the dependent variable
table(donors$donated)
## returns:   
## 0     1 
## 88751  4711

# Build the donation model
donation_model <- glm(donated ~ bad_address + interest_religion + interest_veterans, 
                      data = donors, family = "binomial")

# Summarize the model results
summary(donation_model)

## some of the output:
##Deviance Residuals: 
##    Min       1Q   Median       3Q      Max  
## -0.3480  -0.3192  -0.3192  -0.3192   2.5678  
##
## Coefficients:
##                   Estimate Std. Error  z value Pr(>|z|)    
## (Intercept)       -2.95139    0.01652 -178.664   <2e-16 ***
## bad_address       -0.30780    0.14348   -2.145   0.0319 *  
## interest_religion  0.06724    0.05069    1.327   0.1847    
## interest_veterans  0.11009    0.04676    2.354   0.0186 *  
##
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

# Estimate the donation probability
donors$donation_prob <- predict(donation_model, type = "response")

# Find the donation probability of the average prospect
mean(donors$donated)
#prints: 0.05040551

# Predict a donation if probability of donation is greater than average (0.0504)
donors$donation_pred <- ifelse(donors$donation_prob > 0.0504, 1, 0)

# Calculate the model's accuracy
mean(donors$donated == donors$donation_pred)
#prints:  0.794815
```

#### Calculating ROC Curves and AUC

__Prompt:__  The dataset donors with the column of predicted probabilities, donation_prob ,is already loaded in your workspace.

Load the pROC package.
Create a ROC curve with roc() and the columns of actual and predicted donations. Store the result as ROC.
Use plot() to draw the ROC object. Specify col = "blue" to color the curve blue.
Compute the area under the curve with auc().

```{r}
# Load the pROC package
library

# Create a ROC curve
ROC <- roc(donors$donated, donors$donation_prob) 
## note: [0,1,0,1,0,0] vs [0.04,0.05,0.01,0.04]

# Plot the ROC curve
plot(ROC, col = "blue")

# Calculate the area under the curve (AUC)
auc(ROC)
```

#### Coding categorical features

```{r}
# Convert the wealth rating to a factor (from numbers to strings)
donors$wealth_rating <- factor(donors$wealth_rating, levels = c(0,1,2,3), labels = c("Unknown", "Low","Medium", "High"))

# Use relevel() to change reference category
donors$wealth_rating <- relevel(donors$wealth_rating , ref = "Medium")

# See how our factor coding impacts the model
summary(glm(donated ~ wealth_rating, data = donors, family = "binomial"))
```

#### Handling missing data by impute / imputing

```{r}
# Find the average age among non-missing values
summary(donors$age)

# Impute missing age values with mean(age)
donors$imputed_age <- ifelse(is.na(donors$age),round(mean(donors$age, na.rm = TRUE),2),donors$age)

# Create missing value indicator for age
donors$missing_age <- ifelse(is.na(donors$age),1,0)
```

#### Building a more sophisticated model
 
Background:  One of the best predictors of future giving is a history of recent, frequent, and large gifts. In marketing terms, this is known as R/F/M:Recency, Frequency, and Money. 

Donors that haven't given both recently and frequently may be especially likely to give again; in other words, the combined impact of recency and frequency may be greater than the sum of the separate effects. Because these predictors together have a greater impact on the dependent variable, their joint effect must be modeled as an interaction.

Instructions:

The donors dataset has been loaded for you.

Create a logistic regression model of donated as a function of money plus the interaction of recency and frequency. Use * to add the interaction term.
Examine the model's summary() to confirm the interaction effect was added.
Save the model's predicted probabilities as rfm_prob. Use the predict() function, and remember to set the type argument.
Plot a ROC curve by using the function roc(). Remember, this function takes the column of outcomes and the vector of predictions.
Compute the AUC for the new model with the function auc() and compare performance to the simpler model.

```{r}
# Build a recency, frequency, and money (RFM) model
rfm_model <- glm(donated ~ money + recency * frequency, data = donors, family = "binomial")

# Summarize the RFM model to see how the parameters were coded
summary(rfm_model)

# Compute predicted probabilities for the RFM model
rfm_prob <- predict(rfm_model, type = "response")

# Plot the ROC curve and find AUC for the new model
library(pROC)
ROC <- roc(donors$donated, rfm_prob)
plot(ROC, col = "red")
auc(ROC)
```

#### Automatic feature selection

In the absence of subject-matter expertise, stepwise regression can assist with the search for the most important predictors of the outcome of interest.

In this exercise, you will use a forward stepwise approach to add predictors to the model one-by-one until no additional benefit is seen.

__Instructions:__ The donors dataset has been loaded for you. Use the R formula interface with glm() to specify the base model with no predictors. Set the explanatory variable equal to 1. Use the R formula interface again with glm() to specify the model with all predictors. Apply step() to these models to perform forward stepwise regression. Set the first argument to null_model and set direction = "forward". This might take a while (up to 10 or 15 seconds) as your computer has to fit quite a few different models to perform stepwise selection. Create a vector of predicted probabilities using the predict() function. Plot the ROC curve with roc() and plot() and compute the AUC of the stepwise model with auc()

```{r}
# Specify a null model with no predictors
null_model <- glm(donated ~ 1, data = donors, family = "binomial")

# Specify the full model using all of the potential predictors
full_model <- glm(donated ~ ., data = donors, family = "binomial")

# Use a forward stepwise algorithm to build a parsimonious model
step_model <- step(null_model, scope = list(lower = null_model, upper = full_model), direction = "forward")

# Estimate the stepwise donation probability
step_prob <- predict(step_model, type = "response")

# Plot the ROC of the stepwise model
library(pROC)
ROC <- roc(donors$donated, step_prob)
plot(ROC, col = "red")
auc(ROC)
```




