---
title: "Vectors"
output: html_notebook
---

#### Vectors

`c()` is the method to create a vector. The c in `c()` stands for combine.  Vectors can hold numeric, character, and logical.  Example:

```{r}
numeric_vector <- c(1, 10, 49)
character_vector <- c("a", "b", "c")
boolean_vector <- c(TRUE, FALSE, TRUE)
```

#### Naming pre-existing vectors

Naming vectors are important to keep track of what the data you're holding is for.  For instance, a vector could hold a series of data (e.g. earnings) for each day.  But we need a way to name which data corresponds to which. The code below illustrates the labeling standard.

```{r}
 # Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)
# Assign days as names of poker_vector
names(poker_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
# Assign days as names of roulette_vectors
names(roulette_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
```

#### Naming vectors with reusable labels

In the previous code, we had to manually retype the labels to each variables.  Reusability is key like in any OOL.  

We could refactor the code before into the following code:

```{r}
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector  
names(roulette_vector) <- days_vector
```

#### `sum()`

`sum()` is a very useful method to sum vector components.  Going back to the example in the previous sections, `sum()` can be used in the following way to calculate the net profits from winnings:

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector
# Total winnings with poker
total_poker <- sum(poker_vector)
# Total winnings with roulette
total_roulette <- sum(roulette_vector)
# Total winnings overall
total_week <- total_poker + total_roulette 
# Print out total_week
total_week
```

```{r}

```

#### Indexing in R

The first element is 1, not 0, as in most programming languages.  Let's say we want to access the poker data from Wednesday:

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
# Define a new variable based on a selection
poker_wednesday <- poker_vector[3]
poker_wednesday
```

#### Accessing multiple entries on an array using vector_name[]

Suppose we want to access the poker results for Tuesday, Wednesday and Thursday then save it on another variable:
```{r}
poker_vector <- c(140, -50, 20, -120, 240)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
poker_midweek <- poker_vector[c(2,3,4)]
# Note:  c(2:4) works too!
poker_midweek
```

__Note:__ It might be tempting to use c(2-4) instead of c(2,3,4) because of other programming languages but it does not work in R. c(2:4) will work though! Again, c(2,3,4) is the same as c(2:4)

#### Mean 

Mean can simply be used by using `mean()` method

```{r}
poker_vector <- c(140, -50, 20, -120, 240)
mean(poker_vector)
```

#### Unique R-Thing:  Selection by comparison.

We can compare a vector to an argument using logical comparison operators (e.g. >, !=) and will return a vector of logical operators indicating TRUE or FALSE

```{r}
poker_vector <- c(140, -50, 20, -120, 240)
# Which days did you make money on poker?
selection_vector <- poker_vector > 0
# Print out selection_vector
selection_vector
```

In the script above, it returned a vector with logical values indicating if profit was made with each day.  

#### Filtering using Selection by comparison.  

From the previous example, selection_vector returned an array of logical values.  

In the previous examples, we were able to select elements in a vector in a variety of ways such as poker_vector[1], poker_vector[2,4], poker_vector[3:5].

Using this idea, we can filter our vector by passing it with another vector full of logical values:

```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
# Which days did you make money on poker?
selection_vector <- poker_vector > 0
# Select from poker_vector these days
poker_winning_days <- poker_vector[selection_vector]
poker_winning_days
```

Here, we had a selection_vector with logical values.  And by passing that vector to poker_vector[args] as the argument, it returned the values that meets the desired paramater.  



