---
title: "KNN"
output: html_notebook
---

## Note:  datasets are not here, so code I wrote here are for reference only

#### Task 1:

After several trips with a human behind the wheel, it is time for the self-driving car to attempt the test course alone.

As it begins to drive away, its camera captures a stop sign. Can you apply a kNN classifier to help the car recognize this sign?

__Instruction:__  

The dataset signs is loaded in your workspace along with the dataframe next_sign, which holds the observation you want to classify.

Load the class package.
Create a vector of sign labels to use with kNN by extracting the column sign_type from signs.
Identify the next_sign using the knn() function.
Set the train argument equal to the signs data frame without the first column.
Set the test argument equal to the data frame next_sign.
Use the vector of labels you created as the cl argument.

Solution:
```{r}
# Load the 'class' package
library(class)

# Create a vector of labels
sign_types <- signs$sign_type

# Classify the next sign observed
knn(train = signs[-1], test = next_sign, cl = sign_types)
```


#### Task 2:

To better understand how the knn() function was able to classify the stop sign, it may help to examine the training dataset it used.

Each previously observed street sign was divided into a 4x4 grid, and the red, green, and blue level for each of the 16 center pixels.  

The result is a dataset that records the sign_type as well as 16 x 3 = 48 color properties of each sign.

_Instruction:_ Use the str() function to examine the signs dataset. Use table() to count the number of observations of each sign type by passing it the column containing the labels. Run the provided aggregate() command to see whether the average red level might vary by sign type.

```{r}
# Examine the structure of the signs dataset
str(signs)

# Count the number of signs of each type
table(signs$sign_type)
## RETURNED ##  
## pedestrian      speed       stop 
##        46         49         51

# Check r10's average red level by sign type
aggregate(r10 ~ sign_type, data = signs, mean)
##  sign_type       r10
## RETURNED ##
#1 pedestrian 113.71739
#2      speed  80.63265
#3       stop 132.39216
```

__Result above makes sense:__  stop sign should have a higher percentage of red.


#### A more advanced example:

Classifying a collection of road signs
Now that the autonomous vehicle has successfully stopped on its own, your team feels confident allowing the car to continue the test course.

The test course includes 59 additional road signs divided into three types. At the conclusion of the trial, you are asked to measure the car's overall performance at recognizing these signs.

__Instructions:__  The class package and the dataset signs are already loaded in your workspace. So is the dataframe test_signs, which holds a set of observations you'll test your model on. Classify the test_signs data using knn(). Set train equal to the observations in signs without labels. Use test_signs for the test argument, again without labels. For the cl argument, use the vector of labels provided for you. Use table() to explore the classifier's performance at identifying the three sign types. Create the vector signs_actual by extracting the labels from test_signs. Pass the vector of predictions and the vector of actual signs to table() to cross tabulate them. Compute the overall accuracy of the kNN learner using the mean() function.


```{r}
# Use kNN to identify the test road signs
sign_types <- signs$sign_type
signs_pred <- knn(train = signs[-1], test = test_signs[-1], cl = sign_types)

# Create a confusion matrix of the actual versus predicted values
signs_actual <- test_signs$sign_type
table(signs_pred, signs_actual)

# Compute the accuracy
mean(signs_actual == signs_pred)
```












