---
title: "Factors Section"
output: html_notebook
---

A factor is a data type that can be used to store categorical variables.  

#### Basics on Factors:

Suppose we have a vector for sex and see what the factor method does to the vector:

```{r}
# Sex vector
sex_vector <- c("Male", "Female", "Female", "Male", "Male")

# Convert sex_vector to a factor
factor_sex_vector <- factor(sex_vector)

# Print out factor_sex_vector
factor_sex_vector
```

From the print out above, we can see that `factor_sex_vector` returns an object with the list (M, F, F, M, M) and the levels involved (F, M).

#### Creating ordinal factors

Suppose we want to create a list with ordinal values for temperature such as Low, Med, High where there is a meaning to the order.  

We can do this by passing two parameters:

1. levels, which passes a vector of categories
2. order = TRUE, to indicate that levels is ordinal to the order it was specificied in levels

```{r}
# Temperature vector
temperature_vector <- c('low', 'high', 'mid', 'low', 'low', 'high')
# Factor 
temperature_factor <- factor(temperature_vector, order=TRUE, levels = c('low', 'mid', 'high'))
# Print factor
temperature_factor
```

#### Renaming levels using `levels()`

Suppose in the original factor object, gender were simply F and M.  And now we want a spelled out version as the actual level.  We can use `levels()` to accomplish such feat.

```{r}
# Code to build factor_survey_vector
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
factor_survey_vector

# Specify the spelled out levels of factor_survey_vector
# Important Note:  we specified Female before Male because factors is ordered alphabetically
levels(factor_survey_vector) <- c("Female", "Male")
factor_survey_vector
```

Again, this is worth mentioning again but c("Female", "Male") we used because of the order in the original factor based on alphabetical order when the factors were simply M and F.  

#### Summary through `summary()`

This is self-explanatory.  `summary()` summarizes our factor object. It can also be used to summarize vectors.  Both will be shown in the following code:

```{r}
# Build factor_survey_vector with clean levels
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
levels(factor_survey_vector) <- c("Female", "Male")
factor_survey_vector

# Generate summary for survey_vector
summary(survey_vector)

# Generate summary for factor_survey_vector
summary(factor_survey_vector)
```

#### Comparing levels

We can use comparison operators when `order = TRUE'.  As an example, let's compare the speed of the 2nd and 5th individual in a vector:

```{r}
# Create factor_speed_vector
speed_vector <- c("medium", "slow", "slow", "medium", "fast")
factor_speed_vector <- factor(speed_vector, ordered = TRUE, levels = c("slow", "medium", "fast"))
# Factor value for second data analyst
da2 <- factor_speed_vector[2]
# Factor value for fifth data analyst
da5 <- factor_speed_vector[5]

# Is data analyst 2 faster than data analyst 5?
da2 > da5
```

The script returns a FALSE.  So the 2nd individual is actually slower.  


