---
title: "List Notebook"
output: html_notebook
---

#### Basics

A list in R allows you to gather a variety of objects under one name (that is, the name of the list) in an ordered way. These objects can be matrices, vectors, data frames, even other lists, etc. 

#### Declaring list object through list()

Suppose we want to declare a list with three different objects.  Let us use `list()` to complete this purpose.  

```{r}
# Vector with numerics from 1 up to 10
my_vector <- 1:10 
# Matrix with numerics from 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)
# First 10 elements of the built-in data frame mtcars
my_df <- mtcars[1:10,]

# Construct list with these different elements:
my_list <- list(my_vector, my_matrix, my_df)
my_list
```

Note:  do note that the vector is [[1]], matrix is [[2]] and the data frame is [[3]] in the list we constructed and called.  The bad thing about this is that we can't use the $ notation for List.

#### Manually naming data inside list

We need to use the same method, `list()`, however, we just need to add a new paramater.  The example is below:

```{r}
# Remember the three variables are already declared in the previous example 
# these variables are my_vector my_matrix my_df

# Construct list with these different elements:
my_list2 <- list(vec = my_vector, mat = my_matrix, df = my_df)
my_list2
```

#### Selecting List data

This is just a basic task and the scripts associated with selection list data:

```{r}
# I will be using the same List data from the previous example called my_list2

# Call the entire mat variable inside our list
my_list2$mat

# Using the $ notation, call the second item in the vec variable in the list.  
my_list2$vec[2]

# W/o the $ notation, call the third item in the vec variable in the list.
my_list2[["vec"]][3]
```


#### Adding vector to a list using c()

Weirdly enough `c()` can be used to combine list and another vector into another list.

Suppose we want to add a 'year' variable to our list with a value of 2017 to our my_list2 and then save that new list to another list called my_list3:

```{r}
# Combining list and num variable to another list
my_list3 <- c(my_list2, year = 2017)
# Printing list to verify that info was indeed added
my_list3

#Verify that year variable exist by using $ notation
my_list3$year
```

By examining the printed my_list3 data, year was indeed added which has an assigned value of 2017, which we expect. 
